### 2019-09-15 ic-journal/formal.tex
Minor changes.

HOL is a formal system of typed logical terms. The types are expressions that denote sets (in the
universe $\mathcal{U}$). HOL's type system is much simpler than those
based on dependent types and other type theories. There are four kinds of types in the HOL
logic, as illustrated in Fig. \ref{fig:hol-types} for its BNF
grammar. Noticed that, in HOL the standard atomic types \emph{bool} and \emph{ind}
 denote, respectively, the distinguished two-element set 2 and the
distinguished infinite set $I$.

@@

Notice that, usually the above four axioms are the only axioms allowed
in conventional formalisation projects in HOL: adding new axioms may
either break the logic consistencies or make the whole work less
convincible. The theorem prover (in general) guarantees that,
without adding new axioms, if both the definitions and the statements
of theorems are correct, then any proof which successfully completes must
be correct.

>>>>

\hl{HOL is a formal system of typed logical terms. The types are expressions that denote sets (in the
universe $\mathcal{U}$). HOL's type system is much simpler than those
based on dependent types and other type theories.} There are four kinds of types in the HOL
logic, as illustrated in Fig. \ref{fig:hol-types} for its BNF
grammar. Noticed that, in HOL the standard atomic types \emph{bool} and \emph{ind}
 denote, respectively, the distinguished two-element set 2 and the
distinguished infinite set $I$.

@@

Notice that, usually the above four axioms are the only axioms allowed
in conventional formalisation projects in HOL: adding new axioms may
either break the logic consistencies or make the whole work less
convincible. The theorem prover (in general) guarantees that,
without adding new axioms, if both the definitions and the statements
of theorems are correct, then any proof which successfully completes must
be correct. \hl{This is also the case here: SOS rules of CCS are
\emph{not} axioms but consequences of an inductive relation.}

### 2019-09-15 ic-journal/intro.tex
Minor changes.

\finish{below: ``In contrast to most literature,...'' we need to
  support this with references}

\hlD{We first consider the case of a single equation (or contraction),
that we call the univariate case, then consider the multivariate
  case, where  more than one equation
  (or contraction) is possible.
The univariate  theorems use
$\lambda$-functions to present  CCS equations,
while the multivariate versions require a
more careful and delicate treatment of 
expressions with several 
variables and substitutions acting on them.
In contrast to most literature, we have adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
\emph{process} is a CCS term without equation variables. 
This allows us a smoother move from the univariate case to the multivariate
one. See} Section~\ref{sec:multivariate} for  details.

@@

Section~\ref{ss:ccs} presents basic background materials on CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Section~\ref{s:eq} discusses equations and contractions.
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
Section~\ref{s:for} highlights our formalisation in HOL4 \hlD{for the
univariate case.}
Section~\ref{sec:multivariate} \hlD{describes the move to the
  multivariate case}. 
Finally,  Section~\ref{s:rel} and \ref{s:concl} discuss related work,
conclusions, and a few directions for future work.

>>>>

We first consider the case of a single equation (or contraction),
that we call the \univariate case, then consider the \multivariate
  case, where  more than one equation
  (or contraction) \hl{with multiple equation variables is considered.}
The \univariate theorems use
$\lambda$-functions to present CCS equations,
while the \multivariate versions require a
more careful and delicate treatment of CCS
expressions with \hl{multiple} variables and substitutions acting on them.

\hl{In contrast to modern literature, e.g.}~\cite{Gorrieri:2015jt},
\hl{we have followed Milner's original approach}~\cite{milner1990operational} and adopted the same
type for both CCS equations and processes: those undefined constants
in CCS terms are treated as  (free) equation variables, and a CCS
\emph{process} is a CCS term without equation variables. 
This allows us a smoother move from the \univariate case to the \multivariate
one. See Section~\ref{sec:multivariate} for  details.

@@

Section~\ref{ss:ccs} presents basic background materials on CCS,
including its syntax, operational semantics, bisimilarity and rooted
bisimilarity.
Section~\ref{s:eq} discusses equations and contractions.
Section~\ref{ss:new} presents rooted contraction and the related
unique-solution result for rooted bisimilarity.
Section~\ref{s:for} highlights our formalisation in HOL4 for the
\univariate case.
Section~\ref{sec:multivariate} describes the \hl{extension} to the
  \multivariate case.
Finally,  Section~\ref{s:rel} and \ref{s:concl} discuss related work,
conclusions, and a few directions for future work.

### 2019-09-15 ic-journal/multivariate.tex
Minor changes.

In this section,  \hlD{we discuss the formalisation of the general}  \multivariate
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}).
We chose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
%\hl{This work is entirely new} after~\cite{EPTCS276.10}.
\hlD{The work is an extension of the univariate case
 in the sense that neither the CCS datatype nor
the SOS rules need to be changed.}

@@

\hlD{In this project, we have  followed} Milner~\cite{Mil89}, thus using the same alphabet for
both agent variables and equation variables. 
\hlD{This has allowed a large reuse of the univariate formalisation. For instance, 
the variable
substitution function used in SOS rules [\texttt{REC}] can be also
used for substituting eqaution variables.
However, a lot of care is needed in 
in the proofs of  many
fundamental lemmas, mostly due to the variable capture issues between
equation variables and agent variables.
}

@@

\begin{proof}
% Milner said it can be proven by ``an easy action
% induction''~\cite{milner1990operational}. However this is not true at all.
% %
%To see things more clearly, 
We  prove a stronger result,
i.e.,  the set of free variables decreases through derivations:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \HOLBoundVar{E\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{FV} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLBoundVar{E}\hfill{[TRANS_FV]}
\end{alltt}

>>>>

In this section,  \hlD{we discuss the formalisation of the} \multivariate
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}).
\hlC{This formalisation only supports finitely many
  equations/contractions and equation variables.\footnote{The original theorems
hold for even (uncountably) infinite equations and equation variables.}}
We chose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
\hlD{The work is an extension of the univariate case}~\cite{EPTCS276.10}
 in the sense that neither the CCS datatype nor
the SOS rules need to be changed.

@@

\hlD{In this project, we have followed} Milner~\cite{Mil89}, using the
same alphabet for both agent variables and equation variables.
This has allowed a large reuse of the \univariate formalisation. \hl{For instance, 
the variable
substitution function used in SOS rules [\texttt{REC}] can be also
used for substituting eqaution variables.
However, a lot of care is needed in the proofs of many
fundamental lemmas, mostly due to the variable capture issues between
equation variables and agent variables.}

@@

\begin{proof}
We  prove a stronger result,
i.e.,  the set of free variables decreases through derivations:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E} \HOLBoundVar{u} \HOLBoundVar{E\sp{\prime}}. \HOLBoundVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenImp{}} \HOLConst{FV} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLBoundVar{E}\hfill{[TRANS_FV]}
\end{alltt}

\hlC{Milner said it can be proved by ``an easy action
induction''}~\cite{milner1990operational}.
